---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# framr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/framr)](https://CRAN.R-project.org/package=framr)
<!-- badges: end -->

The goal of `framr` is to make it easier and faster to perform common tasks associated with the [FRAM model](https://framverse.github.io/fram_doc/).

The package currently consists of high-level convenience functions that typically require both updated R/RStudio installations and various FRAM-related files (e.g., Access project databases, Excel files).


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("FRAMverse/framr")
```

Several functions also require the Windows/Office low-level tools in the `RDCOMClient` package, which (as of `r Sys.Date()`) can be installed with:

``` r
#install.packages("RDCOMClient", repos = "http://www.omegahat.net/R")
```


## Example

The function `aeq_mort` quickly generates AEQ'd mortality values for Chinook. 

```{r example, eval=FALSE}
library(framr)

m_hcff <- aeq_mort(
  db = "path/to/ChinookFRAM.mdb",
  runs = 201, #defaults to all RunIDs
  stocks = 31:32, #defaults to all StockIDS 
  drop_t1 = T, #defaults to TRUE for "fishing year" t2:t4
  sum_ages = T #defaults to FALSE for disaggregated returned object
  )

```
